\textcolor{violet}{Reverse Correlation Uncovers More Complete Tinnitus Spectra \vspace{0.25cm}}

### 2023-01-05 paper/short/main_wc995.tex
minor fixes
Wc995 acl (#176)
* Edit paper to under 1000 words, crop gfx

* Cleanup

* Cleanup

* Merge changes from ACL

* minor fixes

* Remove citations

Co-authored-by: Nelson Barnett <nelsonvbarnett@gmail.com>

% The paper headers
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% \lhead{\flogo}
% \fancyhead[L,RO]{\hl{PAPER ID NUMBER}}
\rhead{\textcolor{violet}{\Large\textbf{Technology Letter         }} \thepage}
\pagestyle{fancyplain}

@@

\author{Alec Hoyland, Nelson V. Barnett, Benjamin W. Roop, Danae Alexandrou, Myah Caplan, Jacob Mills, Benjamin Parrell, Divya A. Chari, and Adam C. Lammert\textsuperscript{\textdagger{}}%
\thanks{This work was supported in part by the University of Massachusetts
Center for Clinical and Translation Science.}%
\thanks{A. Hoyland is with the Department of Biomedical Engineering (BME)
at Worcester Polytechnic Institute (WPI), Worcester, MA and with Clarifai, Inc., Wilmington, DE.
N.V. Barnett, M. Caplan, J. Mills and A.C. Lammert are also with BME at WPI.
B.W. Roop is formerly of the Neuroscience Program at WPI, now with  MIT Lincoln Laboratory, Lexington, MA.
D. Alexandrou is with the Stritch School of Medicine at Loyola University Chicago, Chicago, IL.
B. Parrell is with the Department of Communication Sciences and Disorders and the Waisman Center at University of Wisconsin, Madison, WI. D.A. Chari is with the
University of Massachusetts Chan Medical School, Worcester, MA
and the Massachusetts Eye and Ear Infirmary,
Boston, MA.
The authors declare no competing interests.}%
\thanks{\textsuperscript{\textdagger{}} Corresponding Author}}%

@@

\section{Introduction}
% INTRO PARAGRAPH 1/3
\IEEEPARstart{T}{INNITUS}\textemdash{}the perception of sound
in the absence of any corresponding external stimulus\textemdash{}affects up to 50 million people in the U.S. \cite{bhattPrevalenceSeverityExposures2016,nationalinstituteondeafnessandothercommunicationdisordersQuickStatisticsHearing},
a third of whom experience functional cognitive impairment and diminished quality of life
\cite{ahmedImpactTinnitusPerception2017,nondahlImpactTinnitusQuality2007,tegg-quinnImpactTinnitusCognition2016}.
Clinical guidelines for tinnitus management involve targeted exposure to tinnitus-like sounds, for habituation therapy (termed \textit{sound therapy}), or as part of Cognitive Behavioral Therapy \cite{tunkelClinicalPracticeGuideline2014}.
Critically, sound therapy outcomes improve when closely informed by the patient's internal tinnitus experience \cite{davisNeuromonicsTinnitusTreatment2007,landgrebeMethodologicalAspectsClinical2012,nickelOutcomeResearchMusic2005,okamotoListeningTailormadeNotched2010,schaetteAcousticStimulationTreatments2010,steinInhibitioninducedPlasticityTinnitus2015,tassCounteractingTinnitusAcoustic2012,wangStateArtSound2020}.
However, existing strategies for characterizing tinnitus sounds, such as Pitch Matching (PM), are best suited for patients whose tinnitus resembles pure tones (\eg ringing), which may represent only $20-50\%$ of patients \cite{turnerjrAuditoryDysfunctionTinnitus1990,henryAuditoryTestResult2009,ukaegbeTinnitusItsEffect2017}.
There is a need for methods to characterize nontonal (\eg buzzing, roaring) tinnitus \cite{henryTinnitusEpidemiologicPerspective2020,henryMeasurementTinnitus2016,norenaPsychoacousticCharacterizationTinnitus2002,robertsResidualInhibitionFunctions2006}.

% INTRO PARAGRAPH 2/3
Nontonal tinnitus sounds are presumed to be complex and heterogeneous
\cite{vajsakovicPrinciplesMethodsPsychoacoustic2021,henryTinnitusEpidemiologicPerspective2020,meikleTinnitusArchiveArchive2004},
although few characteristics have been firmly established.
Therefore, we base our approach on \textit{reverse correlation} (RC),
an established behavioral method
\cite{ahumadaStimulusFeaturesSignal1971,gosselinSuperstitiousPerceptionsReveal2003,brimijoinInternalRepresentationVowel2013}
for estimating internal perceptual representations that is unconstrained by prior knowledge about the representations themselves.
RC asks subjects to render subjective judgments over random stimuli, being closely related to Wiener theory \cite{wienerNonlinearProblemsRandom1966},
which has inspired ``white-noise'' approaches to system characterization in physiology
\cite{ringachReverseCorrelationNeurophysiology2004,marmarelisWhiteNoiseMethodSystem1978} and engineering \cite{ljungMeasureLackFit1978}.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{experiment_overview.pdf}
  \caption{Diagram of the experimental protocol.
  Subjects listen to a series of random stimuli,
  each preceded by a target sound.
  Subjects compare the stimulus to the target, and respond either ``yes'' or ``no''
  depending on their perceived similarity.
  The recorded stimulus-response pairs are used to form an estimate of the target.}
  \label{fig:experimentdiagram}
\end{figure}

@@

Our results show that RC can accurately reconstruct the PTS of complex, non-tonal, tinnitus-like sounds, capturing the most salient features of the target.
Reconstruction accuracies observed here are below the simulated ideal, which may be attributed to noisy responses universally observed in applications of RC, and which may be mitigated by further optimizing the experimental protocol, stimulus generation, and reconstruction method.
For example, recent approaches to improving RC reconstruction methods can boost efficiency, noise robustness and overall accuracy \cite{roopCompressiveSensingApproach2021,comptonStimulusWhiteningImproves2022a}.
Subjects completed the required number of trials within 10 minutes,
indicating that this procedure could be conducted in a single clinical visit.
RC may therefore be useful as the basis for a clinical assay to characterize a wider variety of tinnitus percepts than currently possible.
Future work will focus on validating this approach in patients suffering from tinnitus.

>>>>

% The paper headers
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\flogo}
\fancyhead[L,RO]{\hl{PAPER ID NUMBER}}
\rhead{\textcolor{violet}{\Large\textbf{Technology Letter         }} \thepage}
\pagestyle{fancyplain}

@@

\author{Alec Hoyland, Nelson V. Barnett, Benjamin W. Roop, Danae Alexandrou, Myah Caplan, Jacob Mills, Benjamin Parrell, Divya A. Chari, and Adam C. Lammert\textsuperscript{\textdagger{}}%
\thanks{This work was supported in part by the University of Massachusetts
Center for Clinical and Translation Science.}%
\thanks{A. Hoyland is with the Department of Biomedical Engineering (BME)
at Worcester Polytechnic Institute (WPI), Worcester, MA and with Clarifai, Inc., Wilmington, DE.
N.V. Barnett, M. Caplan, J. Mills and A.C. Lammert are also with BME at WPI.
B.W. Roop is formerly of the Neuroscience Program at WPI.
D. Alexandrou is with the Stritch School of Medicine at Loyola University Chicago, Chicago, IL.
B. Parrell is with the Department of Communication Sciences and Disorders and the Waisman Center at University of Wisconsin, Madison, WI. D.A. Chari is with the
University of Massachusetts Chan Medical School, Worcester, MA
and the Massachusetts Eye and Ear Infirmary,
Boston, MA.
The authors declare no competing interests.}%
\thanks{\textsuperscript{\textdagger{}} Corresponding Author}}%

@@

\section{Introduction}
% INTRO PARAGRAPH 1/3
\IEEEPARstart{T}{INNITUS}\textemdash{}the perception of sound
in the absence of any corresponding external stimulus\textemdash{}affects up to 50 million people in the U.S. \cite{bhattPrevalenceSeverityExposures2016},
a third of whom experience functional cognitive impairment and diminished quality of life
\cite{nondahlImpactTinnitusQuality2007,tegg-quinnImpactTinnitusCognition2016}.
Clinical guidelines for tinnitus management involve targeted exposure to tinnitus-like sounds, for habituation therapy (termed \textit{sound therapy}), or as part of Cognitive Behavioral Therapy \cite{tunkelClinicalPracticeGuideline2014}.
Critically, sound therapy outcomes improve when closely informed by the patient's internal tinnitus experience \cite{okamotoListeningTailormadeNotched2010,schaetteAcousticStimulationTreatments2010,steinInhibitioninducedPlasticityTinnitus2015}.
However, existing strategies for characterizing tinnitus sounds, such as Pitch Matching (PM), are best suited for patients whose tinnitus resembles pure tones (\eg ringing), which may represent only $20-50\%$ of patients \cite{henryAuditoryTestResult2009,ukaegbeTinnitusItsEffect2017}.
There is a need for methods to characterize nontonal (\eg buzzing, roaring) tinnitus \cite{henryMeasurementTinnitus2016,norenaPsychoacousticCharacterizationTinnitus2002}.

% INTRO PARAGRAPH 2/3
Nontonal tinnitus sounds are presumed to be complex and heterogeneous
\cite{vajsakovicPrinciplesMethodsPsychoacoustic2021,henryTinnitusEpidemiologicPerspective2020},
although few characteristics have been firmly established.
Therefore, we base our approach on \textit{reverse correlation} (RC),
an established behavioral method
\cite{ahumadaStimulusFeaturesSignal1971,gosselinSuperstitiousPerceptionsReveal2003,brimijoinInternalRepresentationVowel2013}
for estimating internal perceptual representations that is unconstrained by prior knowledge about the representations themselves.
RC asks subjects to render subjective judgments over random stimuli, being closely related to Wiener theory,
which has inspired ``white-noise'' approaches to system characterization in physiology
\cite{ringachReverseCorrelationNeurophysiology2004,marmarelisWhiteNoiseMethodSystem1978} and engineering \cite{ljungMeasureLackFit1978}.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{gfx/experiment_overview.pdf}
  \caption{Diagram of the experimental protocol.
  Subjects listen to a series of random stimuli,
  each preceded by a target sound.
  Subjects compare the stimulus to the target, and respond either ``yes'' or ``no''
  depending on their perceived similarity.
  The recorded stimulus-response pairs are used to form an estimate of the target.}
  \label{fig:experimentdiagram}
\end{figure}

@@

Our results show that RC can accurately reconstruct the PTS of complex, non-tonal, tinnitus-like sounds, capturing the most salient features of the target.
Reconstruction accuracies observed here are below the simulated ideal, which may be attributed to noisy responses universally observed in applications of RC, and which may be mitigated by further optimizing the experimental protocol, stimulus generation, and reconstruction method.
For example, recent approaches to improving RC reconstruction methods can boost efficiency, noise robustness and overall accuracy \cite{comptonStimulusWhiteningImproves2022a}.
Subjects completed the required number of trials within 10 minutes,
indicating that this procedure could be conducted in a single clinical visit.
RC may therefore be useful as the basis for a clinical assay to characterize a wider variety of tinnitus percepts than currently possible.
Future work will focus on validating this approach in patients suffering from tinnitus.
